An open-source full 3D electromagnetic modeler for 1D VTI media in Python: empymod

### 2017-06-03 ManuscriptGeophysics2016.tex
Manuscript changes for final submission

\cite{PIER.10.Slob} published analytical frequency- and time-domain solutions
for the diffusive electric field in a VTI half-space, where source and receiver
can be placed anywhere in the half-space. As an example, I use the same model
as \cite{GEO.15.Hunziker} in their Figures~1 and~2: A half-space with
horizontal resistivity $\rho_\mathrm{h} = 1/3\,\Omega$\,m, anisotropy $\lambda
= \sqrt{10}$, and for frequency $f = 0.5\,$Hz. The source is located
horizontally at the origin ($x_\mathrm{s} = y_\mathrm{s} = 0\,$m) at a depth of
150\,m, and the depth of the receivers is 200\,m. The field is calculated on a
regular grid with a spacing of 10\,m. Figure~\ref{fig:analytical} shows the
analytical amplitude and phase results for the first quadrant (the other
quadrants are simply symmetric copies of it).
%
\plot*[btp]{analytical}{width=\textwidth}{Analytical solution for a half-space
  with $\rho_\mathrm{h} = 1/3\,\Omega$\,m, $\lambda = \sqrt{10}$, $f =
  0.5\,$Hz, $z_\mathrm{s} = 150\,$m, and $z_\mathrm{r} = 200\,$m, calculated on
  a regular grid with a spacing of 10 meters; (a) amplitude and (b) phase.}
%
The figure shows the result for x-directed electric source and receiver dipoles
($G^\mathrm{ee}_{xx}$), other configurations yield similar results (the choice
is based on the insight that this is the most interesting of all electric
source to electric receiver fields, as can be seen in \cite{GEO.15.Hunziker},
Figures~1 and~2).

The error of the amplitude is shown in Figure~\ref{fig:onederror-amplitude} for
different settings regarding the Hankel transform: (a) for a 51\,pt QWE with
relative tolerance $\varepsilon_r$ and absolute tolerance $\varepsilon_a$ of
$10^{-12}$ and $10^{-30}$, respectively; (b) for a 21\,pt QWE with
$\varepsilon_r = 10^{-8}$ and $\varepsilon_a = 10^{-30}$; (c) for a 15\,pt QWE
with $\varepsilon_r = 10^{-8}$ and $\varepsilon_a = 10^{-18}$; (d) using the
standard FHT method with the filter Key09-201; (e) using the splined FHT of the
same filter with 40 points per decade; and (f) using the lagged FHT of the same
filter.
%
\plot*[btp]{onederror-amplitude}{width=.92\textwidth}{Error levels for
  different Hankel transform settings, compared with the analytical solution in
  Figure~1. The high precision QWE (a) and the standard FHT (d) have very low
  error levels, generally in the order of $10^{-6}$\,\% or less. The lagged FHT
  has much higher error levels, and the effects of interpolation can be seen
  clearly in the ring-like structure. However, almost the entire error is below
  0.1\,\% ($10^{-1}$\,\%), only the dark black parts have an error of 1\,\% or
  slightly more. Hankel arguments for QWE: {[rel.\ tol., abs. tol., nr of
  pts]}; for splined FHT: {[pts/decade]}.}
%
The results from the high precision QWE (a) and the standard FHT (d) are almost
identical, even though they use completely different Hankel transforms. The
error is, in this case, not due to the method used for the Hankel transform,
but rather to the distinct differences between the analytical solution, which
uses the diffusive approximation, and the numerical code, which calculates the
complete field. If the QWE is calculated for lower tolerance levels, as shown
in (b) and (c), or the FHT is used with interpolation as in (e) and (f),
artefacts seem to arise from the Hankel transform and the interpolation. The
error of the phase is shown in Figure~\ref{fig:onederror-phase}, with the same
conclusion.
%
\plot*[btp]{onederror-phase}{width=.92\textwidth}{Same as
  Figure~\ref{fig:onederror-amplitude}, but for the phase. Interesting is to
  see that interpolation in the wavenumber domain (e) yields slightly different
  patterns than interpolation in space domain (f).}
%

Figure~\ref{fig:filtercomp-amplitude} shows a comparison of the error for all
nine included FHT-filters, for the amplitude (phase is not shown, but is very
similar). They are all very accurate, and by choosing an optimal filter one has
to decide between accuracy and speed, as shorter filters are faster. However,
we have to keep in mind that the accuracy of the filters depends on the model,
the offsets, and the frequencies. If we do not have an analytical solution, as
in this case, we cannot check the accuracy.
%
\plot*[btp]{filtercomp-amplitude}{width=.92\textwidth}{Comparison of the nine
  included FHT filters. All of them have an error mostly far below 1\,\%. For
  other models and other frequencies the result will be different. (Phase is
  not shown but looks very similar.)}
%

@@

Figure~\ref{fig:emmod-HS} shows the error of amplitude and phase for EMmod. On
the left side with a colorscale from $10^0$ to $10^2$\,\% as used in
\cite{GEO.15.Hunziker}, on the right side with a colorscale from $10^{-8}$ to
$10^0$\,\% as used in Figures~\ref{fig:onederror-amplitude}
and~\ref{fig:onederror-phase}. Note that the error calculation in
\cite{GEO.15.Hunziker} was done slightly different. Here I show the relative
error between the analytical result and the result from EMmod, where
\cite{GEO.15.Hunziker} shows the relative error between $\log_{10}$(analytical
result) and $\log_{10}$(result from EMmod).
%
\plot*[btp]{emmod-HS}{width=.7\textwidth}{Error of EMmod for the half-space
  model in Figure~1. On the left side with the same colorscale as in
  \cite{GEO.15.Hunziker}, on the right side with the colorscale as in Figures~2
  and~3.}
%
Figure~\ref{fig:emmod-HS} shows a few interesting points. The responses from
EMmod in this example have significantly less precision than the results from
empymod. This does not mean in any way that EMmod is less precise, as EMmod can
be adjusted to yield much preciser results. However, this would significantly
increase the run time, so it has to be kept in mind for the run time
comparison. The important point is that EMmod does \emph{always} do an
interpolation in the space-frequency domain, by default, a linear
interpolation. No interpolation is used in empymod with QWE or FHT as Hankel
transform, unless one specifies it to speed up the calculations (splined QWE or
splined and lagged FHT options), which will therefore yield preciser results.
This can be seen very nicely in the results. Figures~\ref{fig:emmod-HS} (b) and
(d) show white circles where the result is very precise. These are the offsets
close to those where the fields were actually calculated. The black bands
in-between are the areas where the result was interpolated. It also shows that
the spacing between calculated offsets increases with increasing offsets.  The
error patterns from empymod in Figures~\ref{fig:onederror-amplitude} and
\ref{fig:onederror-phase} show in (a) and (d) a different pattern, displaying
differences between the analytical, diffusive solution and the numerical, full
wavefield solution. It is only for the splined and lagged versions that one
sees the same, circular patterns appearing, which are due to interpolation.

@@

Many 1D CSEM codes use the diffusive approximation that is valid for low
frequencies. The appealing part of the derivation of \cite{GEO.15.Hunziker} is
that it models the complete wavefield, hence it is also valid for high
frequencies and therefore wave-phenomena. As an extreme case,
\cite{GEO.15.Hunziker} show a 1D example for ground-penetrating radar with a
center frequency of 250\,MHz. To do so, they calculate 2048 frequencies in the
range of 1\,MHz to 2048\,MHz for the Fast Fourier transform from frequency to
time domain. Here I only calculated the FFT with 850 frequencies from 1\,MHz to
850\,MHz and then zero-padded up to 2048\,MHz, for EMmod and empymod; tests
have shown that this is sufficient. Figure~\ref{fig:gpr} shows the model with
the survey setup and the four results for EMmod, empymod with QUAD for relative
and absolute tolerance of $10^{-12}$ and $10^{-20}$, respectively, empymod with
a 51\,pt QWE and relative and absolute tolerance of $10^{-8}$ and $10^{-15}$,
respectively, and empymod with FHT with the filter Key09-401, together with the
analytical solution for the arrival of the direct wave (red), the wave
refracted at the surface (yellow), and the wave reflected at the subsurface
interface (green).

EMmod and the QUAD and QWE transforms of empymod yield pretty much the same
result. It is only at later times that some noise can be seen, different in
each method. FHT shows the first arrivals well, however, afterwards the result
becomes extremely noisy.
%
\plot*[btp]{gpr}{width=1\textwidth}{GPR example for (b) EMmod, (c) empymod with
  QUAD for relative and absolute tolerance of $10^{-12}$ and $10^{-20}$,
  respectively, (d) empymod with 51\,pt QWE for relative and absolute tolerance
  of $10^{-8}$ and $10^{-15}$, respectively, and (e) empymod with FHT
  (Key09-401); run times are given in the subplot titles.}
%
Interesting are the calculation times for these three results. EMmod took
roughly 9.5 hours to calculate, empymod with QUAD about 7.8 hours, empymod with
QWE about 4.4 hours, and empymod with FHT under a minute. Note that QWE has
internally a check: If the kernel is too steep within given intervals, it uses
QUAD instead, as QWE cannot handle very steep functions. QWE uses in about 1/3
% (37%)
of the calculation of this example QUAD. Surprising is how good the FHT result
is, given that the filter was designed for CSEM data, hence frequencies 6 to 9
orders of magnitudes smaller and much bigger offsets. In any case,
EMmod/empymod are not optimized for nor intended to model GPR data, and
calculations in time-domain will be much faster and more precise. It is
nevertheless an interesting proof of concept and shows that EMmod/empymod model
the entire EM field. This leaves the question if it would be possible to create
a filter that works for the frequencies and the offsets required for GPR
calculations.

@@

These results regarding speed of calculation do not change the fact that the
advantage of the QWE method is to have an estimate of the error. However, the
filters designed for CSEM problems seem to be very good at solving them, as can
be seen in the low error levels in Figures~\ref{fig:onederror-amplitude}
and~\ref{fig:onederror-phase} or in \cite{GEO.12.Key}.

@@

Figure~\ref{fig:impulse} shows the result of this comparison, where FFTLog with
6\,ms is the fastest method and FFT with 562\,ms the slowest one. Interesting
is to see how many frequencies are used internally, as listed in
Table~\ref{tbl:time}.
%
\tabl[btp]{time}{Run-times for the four different Fourier transforms, and
  frequency range they require. More frequencies does not necessarily mean
  faster nor more precise. All Fourier transform do interpolation either in
  frequency or in time domain.}{
  \centering
  \begin{tabular}{lcccc}
  \toprule
  %
   Method  &  \# freq   & min freq & max freq & time \\
           &            & Hz       & Hz       & ms \\
  %
  \midrule
  %
    FFTLog     &  70& 1.8e-5& 1.4e2 &   6\\
    Sine-filter & 128& 5.3e-7& 5.7e3 &  11\\
    QWE        & 178& 7.9e-5& 6.3e4 & 298\\
    FFT        &  61& 5.0e-5& 5.2e1 & 562\\
  %
  \bottomrule
\end{tabular}}%
%
Even though FFT uses the least frequencies it is the slowest (and least
precise) method. The reason is that it only calculates the response for 61
frequencies, but for the actual FFT it afterwards interpolates them at $2^{20}
= 1\,048\,579$ frequencies to have a regular array from 5e-5\,Hz to 52\,Hz. To
get a better result we would have to get to lower and higher frequencies, but
then the number of required frequencies would even grow bigger. Note that this
example is a difficult one, with source and receiver at the interface between
air and the subsurface. Models where source and receiver are not at the same
depth and not at the air-interface, for instance in marine CSEM where source
and receiver are in the water layer, are generally much easier, in other words,
faster and more precise.
%
\plot*[btp]{impulse}{width=\textwidth}{Comparison of different Fourier
  transforms for an impulse response at the interface between air and a
  half-space of $10\,\Omega\,$m at $6\,$km offset. FFTLog is the fastest, and
  FFT the slowest in this example.}
%

@@

Figure~\ref{fig:codestructure} shows the structure of empymod.%
%
\plot*[btp]{codestructure}{}{Code structure of the module empymod showing the
  different files and the main modeling routines \texttt{bipole},
  \texttt{dipole}, and \texttt{analytical}.}%
%
The important routines for a user who only wants to use the code and is not
interested in the details are located in \texttt{model.py}. The three main
modeling routines are \texttt{bipole}, \texttt{dipole}, and
\texttt{analytical}. With \texttt{bipole} you can calculate frequency- and
time-domain responses for arbitrary oriented, electric or magnetic bipole
sources and receivers of finite length. The routine \texttt{dipole} is
basically a subset of \texttt{bipole}, and is a good starting point if you want
to create your own modeling routine. And \texttt{analytical} is an interface to
the analytical full- and half-space solutions that come with empymod. In
addition there are the routines \texttt{wavenumber} to calculate the
wavenumber-frequency domain result of a dipole, and \texttt{gpr} to calculate
the response for a ground-penetrating radar with a given center frequency. More
modeling routines will hopefully be added in the future.

>>>>

\cite{PIER.10.Slob} published analytical frequency- and time-domain solutions
for the diffusive electric field in a VTI half-space, where source and receiver
can be placed anywhere in the half-space. As an example, I use the same model
as \cite{GEO.15.Hunziker} in their Figures~1 and~2: A half-space with
horizontal resistivity $\rho_\mathrm{h} = 1/3\,\Omega$\,m, anisotropy $\lambda
= \sqrt{10}$, and for frequency $f = 0.5\,$Hz. The source is located
horizontally at the origin ($x_\mathrm{s} = y_\mathrm{s} = 0\,$m) at a depth of
150\,m, and the depth of the receivers is 200\,m. The field is calculated on a
regular grid with a spacing of 10\,m. Figure~\ref{fig:Figure_1} shows the
analytical amplitude and phase results for the first quadrant (the other
quadrants are simply symmetric copies of it).
%
\plot*[btp]{Figure_1}{width=\textwidth}{Analytical solution for a half-space
  with $\rho_\mathrm{h} = 1/3\,\Omega$\,m, $\lambda = \sqrt{10}$, $f =
  0.5\,$Hz, $z_\mathrm{s} = 150\,$m, and $z_\mathrm{r} = 200\,$m, calculated on
  a regular grid with a spacing of 10 meters; (a) amplitude and (b) phase.}
%
The figure shows the result for x-directed electric source and receiver dipoles
($G^\mathrm{ee}_{xx}$), other configurations yield similar results (the choice
is based on the insight that this is the most interesting of all electric
source to electric receiver fields, as can be seen in \cite{GEO.15.Hunziker},
Figures~1 and~2).

The error of the amplitude is shown in Figure~\ref{fig:Figure_2} for
different settings regarding the Hankel transform: (a) for a 51\,pt QWE with
relative tolerance $\varepsilon_r$ and absolute tolerance $\varepsilon_a$ of
$10^{-12}$ and $10^{-30}$, respectively; (b) for a 21\,pt QWE with
$\varepsilon_r = 10^{-8}$ and $\varepsilon_a = 10^{-30}$; (c) for a 15\,pt QWE
with $\varepsilon_r = 10^{-8}$ and $\varepsilon_a = 10^{-18}$; (d) using the
standard FHT method with the filter Key09-201; (e) using the splined FHT of the
same filter with 40 points per decade; and (f) using the lagged FHT of the same
filter.
%
\plot*[btp]{Figure_2}{width=.92\textwidth}{Error levels for
  different Hankel transform settings, compared with the analytical solution in
  Figure~1. The high precision QWE (a) and the standard FHT (d) have very low
  error levels, generally in the order of $10^{-6}$\,\% or less. The lagged FHT
  has much higher error levels, and the effects of interpolation can be seen
  clearly in the ring-like structure. However, almost the entire error is below
  0.1\,\% ($10^{-1}$\,\%), only the dark black parts have an error of 1\,\% or
  slightly more. Hankel arguments for QWE: {[rel.\ tol., abs. tol., nr of
  pts]}; for splined FHT: {[pts/decade]}.}
%
The results from the high precision QWE (a) and the standard FHT (d) are almost
identical, even though they use completely different Hankel transforms. The
error is, in this case, not due to the method used for the Hankel transform,
but rather to the distinct differences between the analytical solution, which
uses the diffusive approximation, and the numerical code, which calculates the
complete field. If the QWE is calculated for lower tolerance levels, as shown
in (b) and (c), or the FHT is used with interpolation as in (e) and (f),
artefacts seem to arise from the Hankel transform and the interpolation. The
error of the phase is shown in Figure~\ref{fig:Figure_3}, with the same
conclusion.
%
\plot*[btp]{Figure_3}{width=.92\textwidth}{Same as
  Figure~\ref{fig:Figure_2}, but for the phase. Interesting is to
  see that interpolation in the wavenumber domain (e) yields slightly different
  patterns than interpolation in space domain (f).}
%

Figure~\ref{fig:Figure_4} shows a comparison of the error for all
nine included FHT-filters, for the amplitude (phase is not shown, but is very
similar). They are all very accurate, and by choosing an optimal filter one has
to decide between accuracy and speed, as shorter filters are faster. However,
we have to keep in mind that the accuracy of the filters depends on the model,
the offsets, and the frequencies. If we do not have an analytical solution, as
in this case, we cannot check the accuracy.
%
\plot*[btp]{Figure_4}{width=.92\textwidth}{Comparison of the nine
  included FHT filters. All of them have an error mostly far below 1\,\%. For
  other models and other frequencies the result will be different. (Phase is
  not shown but looks very similar.)}
%

@@

Figure~\ref{fig:Figure_5} shows the error of amplitude and phase for EMmod. On
the left side with a colorscale from $10^0$ to $10^2$\,\% as used in
\cite{GEO.15.Hunziker}, on the right side with a colorscale from $10^{-8}$ to
$10^0$\,\% as used in Figures~\ref{fig:Figure_2}
and~\ref{fig:Figure_3}. Note that the error calculation in
\cite{GEO.15.Hunziker} was done slightly different. Here I show the relative
error between the analytical result and the result from EMmod, where
\cite{GEO.15.Hunziker} shows the relative error between $\log_{10}$(analytical
result) and $\log_{10}$(result from EMmod).
%
\sideplot[btp]{Figure_5}{width=.7\textwidth}{Error of EMmod for the half-space
  model in Figure~1. On the left side with the same colorscale as in
  \cite{GEO.15.Hunziker}, on the right side with the colorscale as in Figures~2
  and~3.}
%
Figure~\ref{fig:Figure_5} shows a few interesting points. The responses from
EMmod in this example have significantly less precision than the results from
empymod. This does not mean in any way that EMmod is less precise, as EMmod can
be adjusted to yield much preciser results. However, this would significantly
increase the run time, so it has to be kept in mind for the run time
comparison. The important point is that EMmod does \emph{always} do an
interpolation in the space-frequency domain, by default, a linear
interpolation. No interpolation is used in empymod with QWE or FHT as Hankel
transform, unless one specifies it to speed up the calculations (splined QWE or
splined and lagged FHT options), which will therefore yield preciser results.
This can be seen very nicely in the results. Figures~\ref{fig:Figure_5} (b) and
(d) show white circles where the result is very precise. These are the offsets
close to those where the fields were actually calculated. The black bands
in-between are the areas where the result was interpolated. It also shows that
the spacing between calculated offsets increases with increasing offsets.  The
error patterns from empymod in Figures~\ref{fig:Figure_2} and
\ref{fig:Figure_3} show in (a) and (d) a different pattern, displaying
differences between the analytical, diffusive solution and the numerical, full
wavefield solution. It is only for the splined and lagged versions that one
sees the same, circular patterns appearing, which are due to interpolation.

@@

Many 1D CSEM codes use the diffusive approximation that is valid for low
frequencies. The appealing part of the derivation of \cite{GEO.15.Hunziker} is
that it models the complete wavefield, hence it is also valid for high
frequencies and therefore wave-phenomena. As an extreme case,
\cite{GEO.15.Hunziker} show a 1D example for ground-penetrating radar with a
center frequency of 250\,MHz. To do so, they calculate 2048 frequencies in the
range of 1\,MHz to 2048\,MHz for the Fast Fourier transform from frequency to
time domain. Here I only calculated the FFT with 850 frequencies from 1\,MHz to
850\,MHz and then zero-padded up to 2048\,MHz, for EMmod and empymod; tests
have shown that this is sufficient. Figure~\ref{fig:Figure_6} shows the model
with the survey setup and the four results for EMmod, empymod with QUAD for
relative and absolute tolerance of $10^{-12}$ and $10^{-20}$, respectively,
empymod with a 51\,pt QWE and relative and absolute tolerance of $10^{-8}$ and
$10^{-15}$, respectively, and empymod with FHT with the filter Key09-401,
together with the analytical solution for the arrival of the direct wave (red),
the wave refracted at the surface (yellow), and the wave reflected at the
subsurface interface (green).

EMmod and the QUAD and QWE transforms of empymod yield pretty much the same
result. It is only at later times that some noise can be seen, different in
each method. FHT shows the first arrivals well, however, afterwards the result
becomes extremely noisy.
%
\plot*[btp]{Figure_6}{width=1\textwidth}{GPR example for (b) EMmod, (c) empymod
  with QUAD for relative and absolute tolerance of $10^{-12}$ and $10^{-20}$,
  respectively, (d) empymod with 51\,pt QWE for relative and absolute tolerance
  of $10^{-8}$ and $10^{-15}$, respectively, and (e) empymod with FHT
  (Key09-401); run times are given in the subplot titles.}
%
Interesting are the calculation times for these three results. EMmod took
roughly 9.5 hours to calculate, empymod with QUAD about 7.8 hours, empymod with
QWE about 4.4 hours, and empymod with FHT under a minute. Note that QWE has
internally a check: If the kernel is too steep within given intervals, it uses
QUAD instead, as QWE cannot handle very steep functions. QWE uses in about 1/3
% (37%)
of the calculation of this example QUAD. Surprising is how good the FHT result
is, given that the filter was designed for CSEM data, hence frequencies 6 to 9
orders of magnitudes smaller and much bigger offsets. In any case,
EMmod/empymod are not optimized for nor intended to model GPR data, and
calculations in time-domain will be much faster and more precise. It is
nevertheless an interesting proof of concept and shows that EMmod/empymod model
the entire EM field. This leaves the question if it would be possible to create
a filter that works for the frequencies and the offsets required for GPR
calculations.

@@

These results regarding speed of calculation do not change the fact that the
advantage of the QWE method is to have an estimate of the error. However, the
filters designed for CSEM problems seem to be very good at solving them, as can
be seen in the low error levels in Figures~\ref{fig:Figure_2}
and~\ref{fig:Figure_3} or in \cite{GEO.12.Key}.

@@

Figure~\ref{fig:Figure_7} shows the result of this comparison, where FFTLog with
6\,ms is the fastest method and FFT with 562\,ms the slowest one. Interesting
is to see how many frequencies are used internally, as listed in
Table~\ref{tbl:time}.
%
\tabl[btp]{time}{Run-times for the four different Fourier transforms, and
  frequency range they require. More frequencies does not necessarily mean
  faster nor more precise. All Fourier transform do interpolation either in
  frequency or in time domain.}{
  \centering
  \begin{tabular}{lcccc}
  \toprule
  %
   Method  &  \# freq   & min freq & max freq & time \\
           &            & Hz       & Hz       & ms \\
  %
  \midrule
  %
    FFTLog     &  70& 1.8e-5& 1.4e2 &   6\\
    Sine-filter & 128& 5.3e-7& 5.7e3 &  11\\
    QWE        & 178& 7.9e-5& 6.3e4 & 298\\
    FFT        &  61& 5.0e-5& 5.2e1 & 562\\
  %
  \bottomrule
\end{tabular}}%
%
Even though FFT uses the least frequencies it is the slowest (and least
precise) method. The reason is that it only calculates the response for 61
frequencies, but for the actual FFT it afterwards interpolates them at $2^{20}
= 1\,048\,579$ frequencies to have a regular array from 5e-5\,Hz to 52\,Hz. To
get a better result we would have to get to lower and higher frequencies, but
then the number of required frequencies would even grow bigger. Note that this
example is a difficult one, with source and receiver at the interface between
air and the subsurface. Models where source and receiver are not at the same
depth and not at the air-interface, for instance in marine CSEM where source
and receiver are in the water layer, are generally much easier, in other words,
faster and more precise.
%
\plot*[btp]{Figure_7}{width=\textwidth}{Comparison of different Fourier
  transforms for an impulse response at the interface between air and a
  half-space of $10\,\Omega\,$m at $6\,$km offset. FFTLog is the fastest, and
  FFT the slowest in this example.}
%

@@

Figure~\ref{fig:Figure_8} shows the structure of empymod.%
%
\plot*[btp]{Figure_8}{}{Code structure of the module empymod showing the
  different files and the main modeling routines \texttt{bipole},
  \texttt{dipole}, and \texttt{analytical}.}%
%
The important routines for a user who only wants to use the code and is not
interested in the details are located in \texttt{model.py}. The three main
modeling routines are \texttt{bipole}, \texttt{dipole}, and
\texttt{analytical}. With \texttt{bipole} you can calculate frequency- and
time-domain responses for arbitrary oriented, electric or magnetic bipole
sources and receivers of finite length. The routine \texttt{dipole} is
basically a subset of \texttt{bipole}, and is a good starting point if you want
to create your own modeling routine. And \texttt{analytical} is an interface to
the analytical full- and half-space solutions that come with empymod. In
addition there are the routines \texttt{wavenumber} to calculate the
wavenumber-frequency domain result of a dipole, and \texttt{gpr} to calculate
the response for a ground-penetrating radar with a given center frequency. More
modeling routines will hopefully be added in the future.
